\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-11-10 tex/ms.tex
working on paper

    \mathbf{e_4}(\pmb{\theta})

@@

    \mathbf{E_4}(\pmb{\theta})

@@

    \mathbf{e_1}(\pmb{\theta}_r)

@@

    \mathbf{e_2}(\pmb{\theta}_\phi, \mathbf{e_1})

@@

    \mathbf{e_1} \,

@@

    \mathbf{e_3}(\pmb{\theta}_\lambda, \mathbf{e_2})

@@

    \mathbf{e_2} \,

@@

    \mathbf{e_4}(\pmb{\theta}_c, \mathbf{e_3})

@@

    \mathbf{e_3} \,

@@

    \mathbf{E_1}(\pmb{\theta}_r)

@@

    \mathbf{E_2}(\pmb{\theta}_\phi, \mathbf{E_1})

@@

    \mathbf{E_1} \,
    \mathbf{E_1}^\top \,

@@

    \mathbf{E_3}(\pmb{\theta}_\lambda, \mathbf{E_2})

@@

    \mathbf{E_2} \,
    \mathbf{E_2}^\top \,

@@

    \mathbf{E_4}(\pmb{\theta}_c, \mathbf{E_3})

@@

    \mathbf{E_3} \,
    \mathbf{E_3}^\top \,

@@

>>>>

    \mathbf{e}_c

@@

    \mathbf{E}_c

@@

    \mathbf{e}_r

@@

    \mathbf{e}_\phi

@@

    \mathbf{e}_r \,

@@

    \mathbf{e}_\lambda

@@

    \mathbf{e}_\phi \,

@@

    \mathbf{e}_c

@@

    \mathbf{e}_\lambda \,

@@

    \mathbf{E}_r

@@

    \mathbf{E}_\phi

@@

    \mathbf{E}_r \,
    \mathbf{E}_r^\top \,

@@

    \mathbf{E}_\lambda

@@

    \mathbf{E}_\phi \,
    \mathbf{E}_\phi^\top \,

@@

    \mathbf{E}_c

@@

    \mathbf{E}_\lambda \,
    \mathbf{E}_\lambda^\top \,

@@

\section{The Size and Contrast Integrals}
\label{sec:size-contrast}

\section{The Latitude and Longitude Integrals}
\label{sec:lat-lon}
%
Our goal in this section is to compute the integrals...
The Wigner rotation matrix for real spherical harmonics up to degree $l_{\mathrm{max}}$
may be written as the block-diagonal matrix
    \setstackgap{L}{1.25\baselineskip}
    \fixTABwidth{T}
    \mathbf{R} =
    \parenMatrixstack{
    \quad\quad \, \mathbf{R}^0 \, \quad\quad
     &              &              &        &                               \\
     & \mathbf{R}^1 &              &        &                               \\
     &              & \mathbf{R}^2 &        &                               \\
     &              &              & \ddots &                               \\
     &              &              &        & \mathbf{R}^{l_{\mathrm{max}}}
    }\quad,
    \mathbf{R}^l = {\mathbf{U}^l}^{-1} \mathbf{D}^l \mathbf{U}^l
is the Wigner rotation matrix for a single spherical harmonic degree,
    \label{eq:U}
    \setstackgap{L}{1.25\baselineskip}
    \fixTABwidth{T}
    \mathbf{U}^l =
    \frac{1}{\sqrt{2}}
    \parenMatrixstack{
        \quad\quad\, \ddots \, \quad\quad\quad\quad
           &       &        &       &          &    &   &    & \Ddots \\
           & \imag &        &       &          &    &   & 1  &        \\
           &       & \imag  &       &          &    & 1 &    &        \\
           &       &        & \imag &          & 1  &   &    &        \\
           &       &        &       & \sqrt{2} &    &   &    &        \\
           &       &        & \imag &          & -1 &   &    &        \\
           &       & -\imag &       &          &    & 1 &    &        \\
           & \imag &        &       &          &    &   & -1 &        \\
    \Ddots &       &        &       &          &    &   &    & \ddots
    }
describes the transformation from complex to real spherical harmonics,
and $\mathbf{D}$ is the Wigner matrix for complex spherical harmonics,
whose terms are given by the expression
    D_{m,m'}^l(\alpha, \beta, \gamma) = \mathrm{e}^{-\imag m' \alpha}
    d_{m,m'}^l(\beta) \mathrm{e}^{-\imag m \gamma}
where $\alpha$, $\beta$, and $\gamma$ are the Euler angles describing
the rotation in the
$\hat{\mathbf{z}}{-}\hat{\mathbf{y}}{-}\hat{\mathbf{z}}$ convention
and $\imag$ is the imaginary unit. The terms of the $d$-matrix
depend on powers of $\sin\left(\nicefrac{\beta}{2}\right)$
and $\cos\left(\nicefrac{\beta}{2}\right)$
\citep[c.f. Equation C15 in][]{Luger2019}, but it is convenient to use
the half-angle formula to express these terms instead as
    d_{m,m'}^l(\beta) =
    \sum\limits_i c_{m,m',i}^{l}
    \mathrm{sgn}(\sin\beta)^{i}
    (1 - \cos\beta)^{\frac{2l - i}{2}}
    (1 + \cos\beta)^\frac{i}{2}
    c_{m,m',i}^{l} & =
    \resizebox{.75\hsize}{!}{$
            \begin{cases}
                \dfrac{
                \left(-1\right)^{\frac{2l + 3m + m' - i}{2}}
                \sqrt{\left(l - m\right)!
                    \left(l + m\right)!
                    \left(l - m'\right)!
                    \left(l + m'\right)!}
                }{
                2^l
                \left(\frac{i - m - m'}{2}\right)!
                \left(\frac{i + m + m'}{2}\right)!
                \left(\frac{2l - i - m + m'}{2}\right)!
                \left(\frac{2l - i + m - m'}{2}\right)!
                }
                 & m - m' - i \,\, \mathrm{even}
                \\[0.5em]
                0
                 & m - m' - i \,\, \mathrm{odd}
            \end{cases}
        $}
Rotation by
an angle $\phi$ about the $\hat{\mathbf{x}}$ axis may be accomplished by choosing
Euler angles
    \alpha & = \frac{\pi}{2} \nonumber \\
    \beta  & = \phi \nonumber          \\
    \gamma & = -\frac{\pi}{2}
which makes it clear that all terms in
$\mathbf{R}_{\hat{\mathbf{x}}}(\phi)$ are equal to (weighted) sums of powers
of $(1 \pm \cos\phi)$.
Since our goal is to compute integrals of these terms multiplied by
a probability density function, it is convenient to model
$\cos\phi$ as a Beta-distributed variable. As we will see, this
choice will allow us to
analytically compute the first two moments of the distribution of
$\phi$ conditioned on $\pmb{\theta}_\phi$.

%
The Beta distribution in $\cos\phi$ has hyperparameters
%
\begin{align}
    \pmb{\theta}_\phi = \left(
    \alpha \, \, \, \,
    \beta \right)^\top
\end{align}
%
and PDF given by
%
\begin{align}
    \label{eq:cosphi-pdf}
    p \big(\cos\phi \, \big| \, \pmb{\theta}_\phi \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function. The implied distribution for $\phi$
may be computed by a straightforward change of variable:
%
\begin{align}
    \label{eq:phi-pdf}
    p \big(\phi \, \big| \, \pmb{\theta}_\phi \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{2\Gamma(\alpha)\Gamma(\beta)}
    \big|
    \sin\phi
    \big|
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{\beta - 1}
    \quad,
\end{align}
%
for $\phi \in \left[ -\frac{\pi}{2}, \frac{\pi}{2} \right]$.

Since the Wigner matrices are block diagonal, we may evaluate the moments of the
distribution one spherical harmonic degree at a time. To that end, let us
write
%
\begin{align}
    \mathbf{e}_\phi
     & \equiv
    \int
    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
    \mathbf{e}_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \nonumber
    \\
    %
     & =
    %
    \left(
    \mathbf{e}_\phi^0
    \,\,\,
    \mathbf{e}_\phi^1
    \,\,\,
    \mathbf{e}_\phi^2
    \,\,\,
    \cdots
    \,\,\,
    \mathbf{e}_\phi^{l_{\mathrm{max}}}
    \right)^\top
    \quad,
\end{align}
%
where
%
\begin{align}
    \mathbf{e}_\phi^l
     & =
    \int
    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
    \mathbf{e}^l_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \nonumber \\
    %
     & =
    %
    {\mathbf{U}^l}^{-1}
    \pmb{\rho}^l
    \,
    \mathbf{U}^l
    \quad,
\end{align}
%
where we define
%
\begin{align}
    \pmb{\rho}^l
     & \equiv
    \int
    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
    \mathbf{e}^l_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \,
\end{align}
%
with
%
\begin{align}
    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi)
     & =
    \mathbf{D}^l\left(\frac{\pi}{2}, \phi, -\frac{\pi}{2}\right)
    \quad.
\end{align}
%
We may write the terms of $\pmb{\rho}^l$ as
%
\begin{align}
    \rho^l_m
     & =
    \int
    \sum\limits_{m'}
    {D^l_{\hat{\mathbf{x}}}}_{m,m'}(\phi) \,
    {e^l_r}_{m'} \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    %
    \nonumber \\[0.5em]
    %
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    \sum\limits_{m'}
    {e^l_r}_{m'}
    \mathrm{e}^{\frac{\imag \pi}{2}(m - m')}
    \sum\limits_i c_{m,m',i}^{l}
    \,
    q^l_i(\pmb{\theta}_{\phi})
    \quad,
\end{align}
%
where
%
\begin{align}
    q^l_i(\pmb{\theta}_{\phi})
     & =
    \frac{1}{2}
    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
    %
    \mathrm{sgn}(\sin\phi)^{i}
    \big| \sin\phi \big|
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
    (1 + \cos\phi)^\frac{i}{2}
    %
    \,
    \mathrm{d}\phi
    \nonumber \\[0.5em]
    %
     & =
    % %
    % \begin{cases}
    %     \displaystyle\int_{0}^{\frac{\pi}{2}}
    %     %
    %     (\cos\phi)^{\alpha - 1}
    %     (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
    %     (1 + \cos\phi)^\frac{i}{2}
    %     %
    %     \,
    %     \sin\phi
    %     \,
    %     \mathrm{d}\phi
    %     %
    %     & i \,\, \mathrm{even}
    %     %
    %     \\
    %     %
    %     0
    %     %
    %     & i \,\, \mathrm{odd}
    % \end{cases}
    % \nonumber \\[0.5em]
    % & =
    % %
    \begin{cases}
        \displaystyle\int_{0}^{1}
        %
        x^{\alpha - 1}
        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
        (1 + x)^\frac{i}{2}
        %
        \,
        \mathrm{d}x
        %
         & i \,\, \mathrm{even}
        %
        \\
        %
        0
        %
         & i \,\, \mathrm{odd} \quad,
    \end{cases}
\end{align}
%
and in the last line we made use of the transformation $x = \cos\phi$.
%
The integral in the expression above has a closed-form solution in terms
of the hypergeometric function $_2F_1$:
%
\begin{align}
    %
    q^l_i(\pmb{\theta}_{\phi})
     & =
    \begin{cases}
        \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
        \,
        _2F_1\Big(
        -\frac{i}{2},
        \alpha;
        l + \alpha + \beta - \frac{i}{2};
        -1
        \Big)
        %
         & i \,\, \mathrm{even}
        %
        \\
        %
        0
        %
         & i \,\, \mathrm{odd} \quad.
    \end{cases}
\end{align}
